Expected Dependency Pair Match:\\ Predicting {HTER} improvements with expected syntactic structure

### 2009-05-11 mtjournal.tex
roughed out draft of HTER correlations section

\section{Per-document correlations with HTER}
\label{sec:hter1}
In this section, we move from an exploration of the space of possible
DPM metrics to a single chosen DPM-derived metric (EDPM).  Here, we
compare EDPM to other metrics in correlation with document-level HTER
performance.

@@

The GALE 2.5 translation corpus is made up of system translations into
English from three sites.
% MO added:
The three sites all use system combination to integrate results from
multiple systems, some of which are phrase-based and some which may
use syntax on either the source or target side. No system provided
system-generated parses.
%
The corpus being translated comes from
Arabic and Chinese in four genres: \texttt{bc} (broadcast
conversation), \texttt{bn} (broadcast news), \texttt{nw} (newswire),
and \texttt{wb} (web text), with corpus sizes shown in
table~\ref{tab:galestats}.
\begin{table}
  \begin{tabular}{r|rr|rr|rr}
    \hline
     & \multicolumn{2}{c|}{Arabic} & \multicolumn{2}{c|}{Chinese}
     & \multicolumn{2}{c}{Total}\\
     & doc & sent & doc & sent & doc   & sent\\
     \hline
     \texttt{bc}    & 59  & 750 & 56 & 1061 & 115 & 1811\\
     \texttt{bn}    & 63  & 666 & 63 & 620  & 126 & 1286\\
     \texttt{nw}    & 68  & 494 & 70 & 440  & 138 & 934 \\
     \texttt{wb}    & 69  & 683 & 68 & 588  & 137 & 1271\\
     \hline
     Total & 259 & 2593& 257& 2709 & 516 & 5302\\
     \hline
  \end{tabular}
  \caption{Corpus statistics for the GALE 2.5 translation
    corpus.}
  \label{tab:galestats}
\end{table}
The corpus includes one English reference translation $r_i$
\cite{gale08phase2_5references} for each sentence $i$ and a system
translation $t_{i,z}$ for each of the three systems $z$. Additionally,
each of the system translations of each segment $i$ has a
corresponding human-targeted reference aligned at the sentence level,
so we have available the HTER score of each segment
$s_{\textrm{HTER}}(t_{i,z})$ at both the sentence and document level.

@@

In table~\ref{tab:hterperdoc}, we show per-document Pearson's $r$
between $\bar{\mu}$EDPM and $\bar{\mu}$HTER, as well as two other
baselines: $\bar{\mu}$TER and $\bar{\mu}$BLEU$_4$.  
\begin{table}
  \begin{tabular}{r|rrrr|rr|r}
    \hline
    $r$ vs. $\bar{\mu}$HTER & \multicolumn{1}{c}{\texttt{bc}}
    & \multicolumn{1}{c}{\texttt{bn}} &
    \multicolumn{1}{c}{\texttt{nw}} & \multicolumn{1}{c}{\texttt{wb}}
    & \multicolumn{1}{|c}{all Arabic} & \multicolumn{1}{c|}{all Chinese}
    & \multicolumn{1}{c}{all} \\
    \hline
    $\bar{\mu}$TER
    &  0.59 &  \textbf{0.35} &  \textbf{0.47}&  \textit{0.17}
    & \textbf{0.54} & \textbf{0.32} &  0.44\\
    $\bar{\mu}$BLEU
    & -0.42 & \textbf{-0.32} & \textbf{-0.46}&  \textbf{-0.27}
    & -0.42  &  \textbf{-0.33} & -0.37 \\
    $\bar{\mu}$EDPM
    & \textbf{-0.69} & \textbf{-0.39} & \textbf{-0.47} &
    \textbf{-0.27}
    & \textbf{-0.60} & \textbf{-0.39} & \textbf{-0.50} \\
    \hline
  \end{tabular}
  \caption{Per-document correlations of $\bar{\mu}$EDPM and others to
    $\bar{\mu}$HTER, by genre and by source language. Bold numbers are
  within 95\% significance of the best per column; italics indicate
  that the sign of the $r$ value has less than 95\% confidence.}
  \label{tab:hterperdoc}
\end{table}
We see that the EDPM has the highest correlation in each of the
subcorpora created by dividing by genre or by source language, as well
as the corpus as a whole.  Not every difference is above 95\%
confidence, but the trend is the same across all of the
subcorpora. EDPM's strengths are particularly clear in the
unstructured domains (\texttt{wb} and \texttt{bc}).

In table~\ref{tab:hterpersent}, we see a similar story.
\begin{table}
  \begin{tabular}{r|rrrr|rr|r}
    \hline
    $r$ vs. $\bar{\mu}$HTER & \multicolumn{1}{c}{\texttt{bc}}
    & \multicolumn{1}{c}{\texttt{bn}} &
    \multicolumn{1}{c}{\texttt{nw}} & \multicolumn{1}{c}{\texttt{wb}}
    & \multicolumn{1}{|c}{all Arabic} & \multicolumn{1}{c|}{all Chinese}
    & \multicolumn{1}{c}{all} \\
    \hline
    $\bar{\mu}$TER
    &  \textbf{0.44} &  \textbf{0.29} &  \textbf{0.33}&  0.25
    & \textbf{0.44} & 0.25  &  \textbf{0.36}\\
    $\bar{\mu}$BLEU
    & -0.31 & -0.24  & -0.29 &  -0.25
    & -0.31  &  -0.24  & -0.28 \\
    $\bar{\mu}$EDPM
    & \textbf{-0.46} & \textbf{-0.31} & \textbf{-0.34} &
    \textbf{-0.30}
    & \textbf{-0.44} & \textbf{-0.30} & \textbf{-0.37} \\
    \hline
  \end{tabular}
  \caption{Per-sentence, length-weighted correlations of
    $\bar{\mu}$EDPM and others to $\bar{\mu}$HTER, by genre and by
    source language. Bold numbers are within 95\% significance of the
    best per column.}
  \label{tab:hterpersent}
\end{table}

@@

TO DO: NOTE RE: SEGMENT-LEVEL CORRS

%   Note 2: when using per-segment mean-subtracted, r values are smaller
%   and *Arabic* was the language where EDPM did not perform best.
% \begin{verbatim}
%                Ar     Zh     All
%       TER     0.38*  0.04   0.19
%       BLEU_4 -0.10  -0.16  -0.14
%       EDPM   -0.31  -0.19* -0.24*
% \end{verbatim}

@@

>>>>

\section{Correlating EDPM with HTER}
\label{sec:hter1}
In this section, we move from an exploration of the space of possible
DPM metrics to a single chosen DPM-derived metric (EDPM).  Here, we
compare EDPM to other metrics in correlation with document-level and
segment-level HTER performance.

@@

The GALE 2.5 translation corpus is made up of system translations into
English from three sites.
% MO added:
The three sites all use system combination to integrate results from
multiple systems, some of which are phrase-based and some which may
use syntax on either the source or target side. No system provided
system-generated parses.
%
The corpus being translated comes from
Arabic and Chinese in four genres: \texttt{bc} (broadcast
conversation), \texttt{bn} (broadcast news), \texttt{nw} (newswire),
and \texttt{wb} (web text), with corpus sizes shown in
table~\ref{tab:galestats}.
\begin{table}
  \begin{tabular}{r|rr|rr|rr}
    \hline
     & \multicolumn{2}{c|}{Arabic} & \multicolumn{2}{c|}{Chinese}
     & \multicolumn{2}{c}{Total}\\
     & doc & sent & doc & sent & doc   & sent\\
     \hline
     \texttt{bc}    & 59  & 750 & 56 & 1061 & 115 & 1811\\
     \texttt{bn}    & 63  & 666 & 63 & 620  & 126 & 1286\\
     \texttt{nw}    & 68  & 494 & 70 & 440  & 138 & 934 \\
     \texttt{wb}    & 69  & 683 & 68 & 588  & 137 & 1271\\
     \hline
     Total & 259 & 2593& 257& 2709 & 516 & 5302\\
     \hline
  \end{tabular}
  \caption{Corpus statistics for the GALE 2.5 translation
    corpus.}
  \label{tab:galestats}
\end{table}
The corpus includes one English reference translation $r_i$
\cite{gale08phase2_5references} for each sentence $i$ and a system
translation $t_{i,z}$ for each of the three systems $z$. Additionally,
each of the system translations of each segment $i$ has a
corresponding human-targeted reference aligned at the sentence level,
so we have available the HTER score of each segment
HTER$(t_{i,z})$ at both the sentence and document level.

@@

In table~\ref{tab:hterperdoc}, we show per-document Pearson's $r$
between $\bar{\mu}$EDPM and $\bar{\mu}$HTER, as well as two other
baselines: $\bar{\mu}$TER and $\bar{\mu}$BLEU$_4$.  
\begin{subtable}
  \begin{table}
    \begin{tabular}{r|rrrr|rr|r}
      \hline
      $r$ vs. $\bar{\mu}$HTER & \multicolumn{1}{c}{\texttt{bc}}
      & \multicolumn{1}{c}{\texttt{bn}} &
      \multicolumn{1}{c}{\texttt{nw}} & \multicolumn{1}{c}{\texttt{wb}}
      & \multicolumn{1}{|c}{all Arabic} & \multicolumn{1}{c|}{all Chinese}
      & \multicolumn{1}{c}{all} \\
      \hline
      $\bar{\mu}$TER
      &  0.59 &  \textbf{0.35} &  \textbf{0.47}&  \textit{0.17}
      & \textbf{0.54} & \textbf{0.32} &  0.44\\
      $\bar{\mu}$BLEU
      & -0.42 & \textbf{-0.32} & \textbf{-0.46}&  \textbf{-0.27}
      & -0.42  &  \textbf{-0.33} & -0.37 \\
      $\bar{\mu}$EDPM
      & \textbf{-0.69} & \textbf{-0.39} & \textbf{-0.47} &
      \textbf{-0.27}
      & \textbf{-0.60} & \textbf{-0.39} & \textbf{-0.50} \\
      \hline
    \end{tabular}
    \caption{Per-document correlations of $\bar{\mu}$EDPM and others to
      $\bar{\mu}$HTER, by genre and by source language. Bold numbers are
      within 95\% significance of the best per column; italics indicate
      that the sign of the $r$ value has less than 95\% confidence.}
    \label{tab:hterperdoc}
  \end{table}
  \begin{table}
    \begin{tabular}{r|rrrr|rr|r}
      \hline
      $r$ vs. $\bar{\mu}$HTER & \multicolumn{1}{c}{\texttt{bc}}
      & \multicolumn{1}{c}{\texttt{bn}} &
      \multicolumn{1}{c}{\texttt{nw}} & \multicolumn{1}{c}{\texttt{wb}}
      & \multicolumn{1}{|c}{all Arabic} & \multicolumn{1}{c|}{all Chinese}
      & \multicolumn{1}{c}{all} \\
      \hline
      $\bar{\mu}$TER
      &  \textbf{0.44} &  \textbf{0.29} &  \textbf{0.33}&  0.25
      & \textbf{0.44} & 0.25  &  \textbf{0.36}\\
      $\bar{\mu}$BLEU
      & -0.31 & -0.24  & -0.29 &  -0.25
      & -0.31  &  -0.24  & -0.28 \\
      $\bar{\mu}$EDPM
      & \textbf{-0.46} & \textbf{-0.31} & \textbf{-0.34} &
      \textbf{-0.30}
      & \textbf{-0.44} & \textbf{-0.30} & \textbf{-0.37} \\
      \hline
    \end{tabular}
    \caption{Per-sentence, length-weighted correlations of
      $\bar{\mu}$EDPM and others to $\bar{\mu}$HTER, by genre and by
      source language.}
    \label{tab:hterpersent}
  \end{table}
\end{subtable}
We see that the EDPM has the highest correlation in each of the
subcorpora created by dividing by genre or by source language, as well
as the corpus as a whole.  Not every difference is above 95\%
confidence, but the trend is the same across all of the
subcorpora. EDPM's strengths, relative to the other measures, are
particularly clear in the unstructured domains (\texttt{wb} and
\texttt{bc}).

Table~\ref{tab:hterpersent} presents correlations weighted by sentence
length.  Though many of the $r$ values are smaller magnitude in this
correlation, EDPM again has the largest correlation in each category.
TER, however, does quite well with length-weighted sentence
correlation, with $r$ values within 95\% confidence of EDPM scores on
nearly every breakdown.

@@

@@

In the previous section, we observed that TER and EDPM each perform
better than BLEU when considering mean-removed length-weighted
correlations with HTER.  We observe that TER's basic operations
(insertions, deletions, shifts and substitutions) are fundamentally
different from those of EDPM, which is an $F$-style overlap measure
and is seeking to make explicit comparison based on an inferred
syntactic structure.
